---
title: "lab_02"
author: "Kiersten Hacker"
date: "2023-01-07"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Load libraries and establish settings

```{r}
# Turn off scientific notation
options(scipen=999)

# Load the tidyverse. If you have not installed the tidyverse already, remove the # from the next line and run it first.  
# install.packages('tidyverse')
library(tidyverse)
library(lubridate)
```

## Load Data

Let's keep working with the UMD courses and Maryland campaign expense data we used in the pre_lab.

```{r}
umd_courses <- read_rds("data/umd_courses.rds")
maryland_expenses <- read_csv("data/maryland_expenses.csv")
```

## Answer questions

**Q1** Create a new column that divides the expense amounts into the following categories - \$0-\$500, \$500-\$5000, \$5001-\$50000, \$50001\_\$100000, \$100001-\$500000 - giving each an appropriate text value, and count how many expenses are in each group. How would you describe the distribution of expenses in a sentence? Would you change the categories?

**A1. The distribution of expenses is concentrated between the category of $0-$500 and gradually decreases at the dollar amount increases. There are also 35 expenditures that cannot be categorized since they do not have a value in the dataset. I would change the categories to include one that accounts for expenses above $500,000, otherwise there will be a lot of values listed as N/A. **

```{r}
maryland_expenses_divided <- maryland_expenses %>%
  mutate (
    amount_category = case_when(
      amount <= 500 ~ "0 to 500",
      amount > 500 & amount <= 5000 ~ "500 to 5,000",
      amount >5000 & amount <= 50000 ~ "5000 to 50000",
      amount > 50000 & amount <= 100000 ~ "50000 to 100000",
      amount > 100000 & amount <= 500000 ~ "100000 to 500000",
      amount > 500000 ~ "above 500000"
    ))
    
  maryland_expenses_divided %>%
  group_by(amount_category) %>%
  summarise(
    count_amount_category = n()
  ) %>%
  arrange(desc(count_amount_category))
  
```

**Q2.** How many departments have offered the courses with "Climate" in the title and at least one seat? Which department has offered the most? 
**A2. Atmospheric and Oceanic Science has offered the most courses with "Climate" in the title and at least one seat.** 

```{r}
climate_courses <- umd_courses %>% filter(str_detect(title, "Climate") & seats >= 1)
climate_courses %>%
  group_by(department) %>%
  summarise(
    count_courses = n()
  ) %>%
  arrange(desc(count_courses))

```

**Q3.** Of the departments listed in A2, which one is the most interesting to you, and why?
**A3. The most interesting to me is that Germanic Studies has offered 2 courses with "Climate" in their title and at least one seat. I know a lot of the courses in the department study German language and literature, so I am curious which courses address climate and how they do so. Perhaps they study the culture around climate change and Germany's efforts to combat it.** 

**Q4.** Using the Maryland campaign expenses data, change the datatype of the expenditure_date column so that it is a date and not a chr. Then create a new column for the month based on expenditure_date. Then write code that shows total amount spent each month by Wes Moore's committee (you'll need to search the data to find the exact name). Write a sentence describing the pattern of spending over time.
**A4.Wes Moore's committee had the most expenses in October 2022 with other months varying. **

```{r}
maryland_expenses_date <- maryland_expenses %>% mutate(expenditure_date=mdy(expenditure_date))
head(maryland_expenses_date)

maryland_expenses_date %>%
  filter(committee_name == "Moore  Wes For Maryland") %>%
  mutate(month = floor_date(expenditure_date, "month")) %>%
  group_by(month) %>%
  summarise(total_amount = sum(amount)) %>%
  arrange(desc(total_amount))

```

**Q5.** Choose your own adventure: using either the UMD courses or campaign expenses data, ask and answer a question that is different than the previous ones. You can use one of the ones you provided in the pre_lab.
**A5. I originally tried to filter for instructors with an NA value to find a potential story line of how many courses do not have an instructor listed to hold the administration accountable. That didn't work so I decided to look at how many courses per term Keith Yearwood has taught as the professor with the most courses. I mostly looked at this because I was stuck and couldn't get anything else to work. ** 

```{r}
yearwood_courses <- umd_courses %>% filter(instructors == "Keith Yearwood" & seats >= 1)
yearwood_courses %>%
  group_by(department, term) %>%
  summarise(
    count_courses = n()
  ) %>%
  arrange(desc(count_courses))
```
