---
title: "lab_03"
author: "Kiersten Hacker"
date: "2023-02-23"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Load libraries and establish settings

```{r}
# Turn off scientific notation
options(scipen=999)

# Load the tidyverse
library(tidyverse)
# Load janitor
library(janitor)
```

## Load data and clean up the column names

Load the two datasets we worked with in the pre_lab: the Maryland payments to vendors and Maryland grants and loans and, using janitor, change the columns so that they are lower-cased and any spaces are replaced with underscores.

```{r}
# Load in the Maryland state agency payments data, making sure that it correctly sets the datatypes using all of the rows

payments <- read_csv("data/State_of_Maryland_Payments_Data__FY2008_to_FY2023.csv", guess_max=322138)
cleaned_payments <- payments %>%
  clean_names()

# Load in the Maryland grants and loan data

md_grants_loans <- read_csv("data/State_of_Maryland_Grant_and_Loan_Data__FY2009_to_FY2022.csv")
cleaned_md_grants_loans <- md_grants_loans %>%
  clean_names()
```

## Answer questions

Q1: You want to know how much money state agencies have paid to the Prince George's County School system, but to do that you'll need to find all of the variations of how the school system might be listed. Write code to find rows where the vendor name contains the county's school system, starting with a broad search and then narrowing it down. When you have found as many as possible, save the result to a new dataframe called `pg_schools`. How many records did you find?

A1: I found 64 records of payments to the Prince George's County school system. I did a search of the payments data and found 8 ways the Prince George's County school system could appear. From there I filtered to create a new dataframe that showed the entries for each of those 7 ways the data could appear. There are other variations I could have missed, but these are what I found. Some included the apostrophe in "George's" and others did not. I included anything that had to do with the Prince George's County Board of Education or Prince George's County Public Schools in all of their different spellings and abbreviations of PG County.

```{r}
pg_schools<-cleaned_payments %>% filter(vendor_name == "PRINCE GEORGES COUNTY PUBLIC SCHOOL" | vendor_name == "BOARD OF EDUCATION OF PRINCE GEORGES COUNTY MD" | vendor_name ==	"BOARD OF EDUCATION OF PRINCE GEORGE'S COUNTY" | vendor_name == "PRINCE GEORGES PUBLIC SCHOOLS" | vendor_name == "PRINCE GEORGES COUNTY SCHOOL" | vendor_name == "PG COUNTY PUBLIC SCHOOL" | vendor_name == "BD OF EDUCATION OF PG CO")
```

Q2: Using that `pg_schools` dataframe, write code to calculate the total amount of payments to the Prince George's County School system for each fiscal year. Which fiscal year had the largest total amount? What do you notice about the years listed?

A2: Fiscal year 2022 had the largest total amount. I notice that the years almost go in chronological order as their amount descends, with the exception of 2015 which had a higher amount than 2018. 2015 sticks out as an interesting outlier and I wonder why the total does not necessarily fit with the trend of each year having a higher total than the previous one in order. Also, fiscal year 2023 currently has the lowest total amount, but I presume that is because the fiscal year is still ongoing and has not come to a close yet.

```{r}
pg_schools %>%
  group_by(fiscal_year) %>%
  summarise(
    total_amount = sum(amount)
  ) %>%
  arrange(desc(total_amount))
```
Q3. You're interested in examining the vendors paid for capital projects at the University of Maryland. Search the payments data to find those payments, then write code to produce a list of each vendor and the total amount it received. Look carefully at that list - what cleaning will it require to ensure that you have only one row per vendor and zip code? Provide at least three examples of vendor names that are similar and should be combined into one. Describe how you might clean that data so that any nearly identical vendors would look identical - what would you change and how?

A3. It will require cleaning to identify dupes, so you could use the get dupes function and then the clean names function to get rid of the duplicates from the dataset. The zip codes could also be cleaned to make sure they are all 5 digits and they are not duplicates. If we clean the names that are similar but truly are the same company and make them into one row, we can assure that there is one row per each vendor. We could rename the cloumns and fix the data using mutate. A MORTON THOMAS & ASSOC INC and A MORTON THOMAS & ASSOCIATES INC are similar and should be combined into one. So should A&I INC and A&I INC. as well as AEROSOL MONITORING & ANALYSIS INC and AERSOL MONITORING & ANALYSIS INC. To change them to be identical, I would make a new column called renamed or something along those lines using mutate and case when so when a company is misspelled, it will be renamed to the right one combining the two into the new column under the same zip code. 

```{r}
cleaned_payments %>%
  filter(agency_name == "CAPITAL PROJECTS- UNIVERSITY OF MARYLAND") %>%
  group_by(vendor_name) %>%
  summarise(total_amount = sum(amount)) %>%
  arrange(vendor_name)
```

Q4: You want to understand how much state grant money went to stem cell research projects each fiscal year. Search the data to find grants that fit that description. Once you find which columns contain the information you want, your code should standardize the data needed to find just the grants for stem cell research in a single filter. Then create a total amount of money for each fiscal year. Write a sentence summarizing how stem cell research funding has changed over time.

A4: Stem cell research funding has seemed to decrease over time since 2009 when it reached its highest at 14,643,721 dollars. There is a brief period from 2010 to 2014 where it seems to alternate between funds increasing in a year and decresing from the year before, then after 2014 funding decreases until a jump in 2019 where the total was 7,799,987 dollars.

```{r}
cleaned_md_grants_loans %>% 
  filter(str_detect(description, "Stem Cell"), category == "Grant") %>%
	group_by(fiscal_year) %>%
  summarise(total_amount = sum(amount)) %>%
  arrange(desc(total_amount))

```
