---
title: "lab_11"
author: "Kiersten Hacker"
date: "2023-04-27"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## You will need

* Our usual libraries for working with data, plus rvest.

## Load libraries and establish settings

**Task** Create a codeblock and load appropriate packages and settings for this lab.
```{r}
library(rvest)
library(tidyverse)
library(janitor)
library(lubridate)
```


Let's get to scraping.

## Questions

**Q1**. Scrape the listing of available Maryland state grants at https://grants.maryland.gov/Pages/StateGrants.aspx into a dataframe. You should have three columns, one of which is a date, so make sure the date column has a date datatype. Then, write code to count the number of grants opportunities offered by each organization listed in your dataframe, showing the organization with the most grant opportunities first. Which state agency has the most?

**A1: The Maryland Energy Administration has the most grants available given the data that we have. However, I am wondering if the data is an accurate representation. Since the website says post a grant, I am not sure if agencies are posting their own grants, which could mean that not all available grants are being reported on the website. Further, some of the grants on the website, like the Game Changers grant, do not have funding anymore. Also, in the dataframe, there are duplicates of some of the agencies due to abbreviations and variations in the name of the same agency, throwing off the numbers and making them unreliable at first glance.** 

```{r}
grants_url <- "https://grants.maryland.gov/Pages/StateGrants.aspx"

results <- grants_url %>%
  read_html() %>%
  html_table()

results <- results[[1]] %>%
  clean_names() %>%
  mutate(due_date = mdy(due_date))

results %>%
  group_by (organization) %>%
  summarise (
    count_grant_title = n()
  ) %>%
  arrange(desc(count_grant_title))

```

**Q2** Next, let's scrape the list of press releases from Maryland's Office of the Public Defender, https://www.opd.state.md.us/press-releases. This isn't a table, so you'll need to use `html_elements()` and your browser's inspector and do some clean up on the results. The result should be a dataframe with two columns that contain the date and title, and the date column should have a date datatype. The challenge here is figuring out how to isolate the releases.

When you finish scraping into a dataframe, write code to find the press releases that have the word "police" in the title. How many are there and when was the most recent one?

**A2: There are only nine releases containing the word "police" in the title and the latest one came from June 2021. There is no way that nothing significant has happened involving the police in the last two years, so this is really interesting because it shows the lack of transparency and communication.** 

```{r}
opd_url <- "https://www.opd.state.md.us/press-releases"

opd_results <- opd_url %>%
  read_html()

opd_results %>% html_elements('p') %>% html_text() %>% as_tibble()

releases <- opd_results %>% html_elements('a.wixui-rich-text__text') %>% html_text() %>% as_tibble() %>% separate(value, c('date', 'title'), sep=":") %>% drop_na(title) %>% mutate(title = str_squish(title), date = mdy(date))

releases %>%
  filter(str_detect(title, "Police")) %>%
   arrange(desc(date))
```

**Q3** Sen. Ben Cardin, D-Maryland, has posted hundreds of press releases at https://www.cardin.senate.gov/?post_type=press-releases. It would be great to have all of them in a dataframe that has the following columns: date, title and url.

To do this, you will need to scrape the page's html and save that to a variable, and _then_ extract the dates, titles and urls into _separate_ dataframes using html_elements(). And remember how we turn a list into a dataframe. The function `html_text()` pulls out the contents of a tag, but for urls we want the HTML attribute. Rvest gives you a way to extract the URL from a link; google to find out what it is.

At the end, you'll have three dataframes that you want to combine into a single dataframe. When we want to combine the rows of identical dataframes, we used `bind_rows()`. If you were combining columns instead of rows, there's a similar function. Find out what it is and use it to put all of the dataframes together into a single one.

When you're done, rename the columns so they make sense, then make sure the date column is an actual date.

Finally, tell me what questions you could ask of this data. Be creative.

**A3: Do his press releases have a correlation with how well his ratings are? Does posting a lot of releases and taking part in such open communication lead him to gain more votes? What topics does he frequently post about? How many are news about a recent activity he partook in and how many are actually him addressing issues? Were the releases more frequent in some years compared to others, like for example election years? What are his most addressed topics, do they appeal to a certain narrative or group?** 

```{r}
cardin_url <- "https://www.cardin.senate.gov/?post_type=press-releases"

##Date
cardin_results_date <- cardin_url %>%
  read_html()

cardin_results_date %>% html_elements('h5') %>% html_text() %>% as_tibble()

cardin_date <- cardin_results_date %>% html_elements('h5.customBlog_item__date') %>% html_text() %>% as_tibble() 

cardin_date <- cardin_date %>%
  rename(date = value) %>%
  mutate(date = mdy(date))

##Title
cardin_results_title <- cardin_url %>%
  read_html()

cardin_results_title %>% html_elements('h3') %>% html_text() %>% as_tibble()

cardin_title <- cardin_results_title %>% html_elements('h3.customBlog_item__title') %>% html_text() %>% as_tibble() %>% mutate(value = str_squish(value))

##URLs
cardin_results_urls <- cardin_url %>%
  read_html()

cardin_results_urls %>% html_elements('a') %>% html_attr("href") %>% as_tibble()

cardin_urls <- cardin_results_urls %>% html_elements('a.customBlog_item__readMore') %>% html_attr("href") %>% as_tibble() 

##Combine dataframes
cardin_releases <- bind_cols(list(cardin_date, cardin_title, cardin_urls))

cardin_releases <- cardin_releases %>%
  rename(title = value...2, url = value...3)
```
