---
title: "lab_05"
author: "Kiersten Hacker"
date: "2023-03-09"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## You will need

- Tabula

## Load libraries and establish settings

```{r}
# Turn off scientific notation
options(scipen=999)

# Load the tidyverse.
library(tidyverse)
library(janitor)
library(lubridate)
library(knitr)
```

## Get Our PDF

We'll be working with the [911 overdose calls from Baltimore County](https://drive.google.com/file/d/1qkYuojGF_6WKFr5aNQxmewDzcKyOiJFr/view?usp=share_link). You'll want to download it to a place you'll remember (like your Downloads folder, or the labs folder in your repository). The goal is to extract the tables within it, export that to a CSV file, load it into RStudio and ask some questions.

## Extract Data from PDF Using Tabula

Start Tabula, then go to http://127.0.0.1:8080/ in your browser. Click the "Browse" button and find the PDF file and click "open", and then click the "Import button" in Tabula. This will take a few seconds or longer.

This PDF has a single table spread over multiple pages to extract. We're going to make a single dataframe from this table, exporting it to a CSV file that you will load into R. In Tabula, highlight the table and click the "Preview & Export Extracted Data" button. You may want to play with including or excluding the column headers - YOU SHOULD HAVE FIVE COLUMNS OF DATA.

Save the CSV (it should be called `tabula-Baltimore County; Carey, Samantha log OD.csv` by default) to your lab_05/data folder.

From there, you will need to read in the data, and add or fix headers if necessary. You can choose to include the headers from the PDF in your exported CSV files OR to exclude them and add them when importing. `read_csv` allows us to do this ([and more](https://readr.tidyverse.org/reference/read_delim.html)).

## Load and clean up the data in R

You will need to read in and clean up the data so that it can be used for analysis. By "clean" I mean the column headers should not contain spaces and they should have meaningful names, not "x1" or something similar. How you do that is up to you, but you can use select() with or without the minus sign to include or exclude certain columns. You also can use the `rename` function to, well, rename columns. Importantly, you'll need to ensure that any columns containing a date actually have a date datatype. Our friend `lubridate` can help with this.

```{r}
 baltimore_county_od_fixed <- read_csv("data/tabula-Baltimore_County_log_OD.csv", col_names = FALSE) %>% clean_names() %>%
  rename(date= x1, time = x2, case_number = x3, event_type = x4, location = x5) %>%
mutate(date=mdy(date))
glimpse(baltimore_county_od_fixed)

```

## Answer questions

Q1. Write code to generate the number of calls that occurred on each date. Which date in 2022 had the most overdose calls, and how many? Look at the total number of rows in your result and explore the range of dates - based on your result, do you believe there are any days with no overdose calls at all? Explain why or why not.

A1. July 14, 2022 had the most overdose calls in 2022 with 23 calls. There are 329 rows which leads me to think that there are some dates where there were no drug overdose calls since there are 365 days in a year. 

```{r}
baltimore_county_od_fixed %>%
  filter(str_detect(date, "2022")) %>%
  group_by(date) %>%
   summarise(
    count_case_number = n()
  ) %>%
  arrange(desc(count_case_number))
```

Q2. You want to understand if there's a pattern in the day of the week that overdose calls are made. Add a column to your dataframe that displays what day of the week each date represents. You should search for how to do that using lubridate. Then write code to calculate the number of calls for each day of the week, and add a column to that result that calculates the percentage of all calls that occurred on each day of the week (so you want a dataframe with the day of the week, total number of calls and the percentage of calls on that day out of the total number of all calls). Describe your findings to me.

A2. It seems like Friday and Saturday have higher/the highest percentage of calls compared to the other days of the week. This raises the question about what is happening on the weekends and where these calls are taking place. Perhaps more action needs to be taken in nightlife areas or more investigation needs to be done.

```{r}
baltimore_county_od_fixed_days <- baltimore_county_od_fixed %>%
  mutate(weekday = wday(date, label = TRUE)) 

baltimore_county_od_fixed_days_total <- baltimore_county_od_fixed_days %>%  
group_by(weekday) %>%
   summarise(
    count_case_number = n()
  ) %>%
  mutate(pct_calls = count_case_number/4001*100 )

```

Q3. Now let's look at locations. Which ones have the most calls? How would you describe them (feel free to search for more information on them)? Is there anything about the structure of the original data that might make you less confident in the counts by location or date?

A3. The location with the most calls appears to be an intersection which prompts me to wonder if there might be a place near by where people gather to do drugs or if maybe there have been traffic stops. I feel like it would be more likely for drug users to just use the intersection as a travel route if they are coming from a near by place, but it is certainly interesting. Some of the other calls came from police precincts or other intersections or a house that had a high number of calls. In the original data sometimes there are multiple calls for the same location at the same time. I wonder if this may be multiple officers responding or officers calling for backup. This could skew the counts and it is an aspect of the data that certainly raises some questions.

```{r}
baltimore_county_od_fixed_days %>%
  group_by(location) %>%
   summarise(
    count_case_number = n()
  ) %>%
  arrange(desc(count_case_number))
```

Q4. What's the best story idea or question you've seen as a result of the work you've done in this lab?

A4. The biggest questions I've had going through this lab is about the location of the drug overdose calls. I think it would be worth investigating where specifically and when the most calls happen to a solution could be put in place. Maybe there is a patter of houses or intersections that are frequent problem areas and once that is found precisely, we could figure out why that is. Why do a larger percentage of calls happen on a Friday and Saturday? Are there more police? Are people letting loose? Why are so many calls connected to one intersection? Is there a location near by that drug users frequent? Is that a place of transit for people escaping the cops? How can we better serve these areas where drug overdose calls are more prevelant and what are the underlying social causes for this?
