---
title: "lab_01"
author: "Kiersten Hacker"
date: "2023-01-05"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
# Turn off scientific notation
options(scipen=999)
```

## About this lab

To complete this lab, you need to: \* run existing code as directed (look for **Task**). \* modify existing code as directed (look for **Task**). \* write code in empty codeblocks provided to answer questions included (look for **Q**). \* write out the answer in the form of a complete sentence in the space given (look for **A**). When you are finished, commit changes and push to your personal GitHub repo, then submit the URL to this document on ELMS.

Remember to add any new verbs or concepts to your class reference notebook!

## Load libraries and establish settings

**Task** Run the codeblock below to load the Tidyverse. To run the codeblock, hit the little green play button in the upper right corner of the gray area (the codeblock) below.

```{r}
# Load the tidyverse. If you have not installed the tidyverse already, remove the # from the next line and run it first.  
# install.packages('tidyverse')
library(tidyverse)
```

## Load Data

We're going to load the UMD courses data.

**Task** Run the codeblock below to load the data, which is contained in an RDS file.

```{r}
umd_courses <- read_rds("lab_01.rds")
```

## Answer questions

**Q1** Which department has the highest mean (average) number of seats for its classes? You need to show the department with the highest average first.
**A1 Engineering, Cooperative Education has the highest average number of seats for its classes with 314.285714 seats.** 

```{r}
umd_courses %>%
  group_by(department) %>%
  summarise(
    mean_seats = mean(seats),
  ) %>%
  arrange(desc(mean_seats))
```

**Q2** Find out the greatest number of times any individual class has been offered in our dataset. You must show the greatest number of classes first.
**A2 The greatest number of times an individual course has been offered is 21 times. There are 155 classes that have been offered 21 times.** 

```{r}
umd_courses %>%
  group_by(id) %>%
  summarise (
    count_terms = n()
  ) %>%
  arrange(desc(count_terms))
```

**Q3** Count the number of courses for each instructor and department. Which instructor has taught the most? 
**A3 Keith Yearwood has taught the most courses at 58 courses in the Geological Science Department.** 

```{r}
umd_courses %>%
  group_by(department, instructors) %>%
  summarise (
    count_id = n()
  ) %>%
  arrange(desc(count_id))
```

**Q4** Given the results for Q3, what would have made it easier to find the answer? You don't have to know _how_ to accomplish this; I want you to describe in words what would make the answer easier to see.
**A4 It would be easier if we were able to define that we did not want rows returned that contained null values for the instructor column. It would also be helpful if we could see if the number of terms they have taught to see if it would make sense with the number of classes they have taught. For example, someone could have retired but still be listed as the instructor until it was updated. Another thing to look for would be cross-listed courses because an instructor could be teaching only one class, but because it is cross-listed perhaps it would show that they are teaching two courses. I think it would be helpful to also see how many terms they have taught the class.**

**Q5** Take one of the questions you listed in the final question of the pre_lab and try to answer it here in code. If you don't know how to fully answer it, write what code you can and describe in words what else you would need to do. If you got an answer, is it interesting?
**A5 I tried to answer the question of which departments offer the most courses and seats per term and how that has changed per term. I was able to get the answer for the most part, but I could not see which department had the most seats offered as well since it arranged by the total number of courses first.**

```{r}
umd_courses %>%
  group_by(term, department) %>%
  summarise (
    count_id = n(),
    total_seats = sum(seats)
  ) %>%
  arrange(desc(count_id), desc(total_seats))
```
