---
title: "lab_07"
author: "Kiersten Hacker"
date: "2023-03-30"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## You will need

-   A Census API key

## Load libraries and establish settings

You'll need to load two packages for this: the tidyverse and tidycensus.

**Task** load these two packages

```{r}
# Turn off scientific notation
options(scipen=999)

# Load the tidyverse.
library(tidyverse)
library(tidycensus)
```

## Setup Census API

You'll need your Census API Key:

**Task** set your API Key if you haven't already. You won't need to install or overwrite it.

```{r echo=FALSE}
census_api_key("484990e06c3b914605047b549cffe5ba654bdbb1", install=TRUE)
```

## Load and view ACS variables

You'll also need to load the list of American Community Survey variables from the 2021 5-year estimates so you can identify the codes you need:

**Task** load the variables from the 2021 5-year ACS (which is called "acs5") and save them to a dataframe

```{r}
acs <- load_variables(2021, "acs5", cache = TRUE)
```

## Answer questions

**Q1** What is the Maryland jurisdiction with the lowest median age, according to the ACS 2021 5-year estimates? You will need to find the variable for median age first. 
**A1: Baltimore City has the lowest median age according to the ACS 2021 5-year estimates. This is bringing me back to the previous work we did with drug overdose calls that showed how young Baltimore City is.** 

```{r} 
md_median_age <- get_acs(geography = "county",
              variables = c(medage = "B01002_001"),
              state = "MD",
              year = 2021)
md_median_age %>% 
  arrange(estimate)
```

**Q2** Which Maryland counties have a median age of at least 50? You can use the dataframe you just made to answer this, but you must write code to display only those counties with a median age of at least 50. 
**A2: Talbot County and Worcester County both have a median age of at least with with Talbot County at 51 and Worcester County at 50.4.** 

```{r}
md_median_age %>%
  filter(estimate>= 50)
```

**Q3** We're interested in knowing more about Montgomery County zip codes where overdose calls are coming from and the demographic characteristics they possess. In particular, we want to calculate a rate of calls per 1,000 population. To do this, you'll need to:

1. Load the Montgomery County 911 overdose calls data.
2. Create a new dataframe with the total number of calls for each zip code.
3. Find the ACS variable pertaining to the total 18 and up population (there are several choices here; you want the simplest one).
4. Get from the ACS data from all Maryland zip codes using that variable.
5. Join the ACS data to the Montgomery County zip code totals dataframe and rename the estimate column to a meaningful name you've given the ACS variable.
6. Add a column to the joined dataframe that calculates a rate of calls per 1,000 population.

Which zip code has the highest rate of calls? Which has the lowest?

**A3: The zip code 20877 has the highest rate of calls and the zip code 20707 has the lowest rate of calls. The zip code 20889 had one call but all other information was NA due to the fact that the location is Walter Reed Hospital so since there are no permanent residents there the census does not have data for it even though that is what ACS was calling upon.** 

```{r}
montgomery_2022_overdoses <- read_csv("data/montgomery_2022_overdoses.csv")

montgomery_2022_overdoses_zip <- montgomery_2022_overdoses %>%
  group_by (zip) %>%
  summarise (
  count_close_type = n ()
) %>%
  mutate(zip = as.character(zip))

md_zip_codes <- get_acs(geography="zcta", variables = "B09021_001", state='MD', year=2019)

montgomery_2022_overdoses_zip_with_acs <- montgomery_2022_overdoses_zip %>% left_join(md_zip_codes, by=c('zip'='GEOID')) %>%
  rename(pop_over_18 = estimate)

montgomery_2022_overdoses_zip_with_acs %>%
  mutate(per_1000_calls = count_close_type/pop_over_18*1000) %>%
  arrange(desc(per_1000_calls))

```

**Q4** Using [CensusReporter.org](https://censusreporter.org/), examine the zip codes with the highest and lowest rates and compare and contrast their demographic profiles. Is there a story here?

**A4: The zip code with the highest rate of calls is in Gaithersburg which is 42% percent Hispanic, 21% white, 18% Black, 16% Asian and 3% two or more races. The median household income is $77,264 and the median age is 37.6 years old. About 15.4% live under the poverty line. The zip code with the lowest rate of calls is in Laurel which is borders both Montgomery and Prince George's counties. Laurel is 45% Black, 22% white, 11% Asian, 16% Hispanic, 6% two or more races and 1% other. The median income is $91,315 and about 8.4% of people live under the poverty line. The median age is 38. The story here could address the difference in median household income, perhaps wealthier zip codes have fewer overdose calls. I think a story line could come from the fact that there is a larger Hispanic community in the Gaithersburg zip code and a larer number of people living under the poverty line. We could see if there is a correlation or causation relationship between ethnicity and income and the number of overdose calls and look into the ways Montgomery County may be failing the neighborhood.** 

**Q5** Go back to your joined dataframe from Q3 and add two more columns representing the minimum and maximum populations based on the estimate and margin of error, then calculate per capita rates for each. Do you get the same two zip codes for the highest and lowest as you did before? Does that give you more or less confidence in your answers to Q3?

**A5: I got the same two zip codes for the highest and lowest rates which gives me more confidence in my answer to Q3 since the numbers are verified twice and would be accurate for a story.** 

```{r}
montgomery_2022_overdoses_zip_with_acs <- montgomery_2022_overdoses_zip_with_acs %>%
  mutate(min = pop_over_18-moe, max = pop_over_18+moe)

montgomery_2022_overdoses_zip_with_acs %>%
  mutate(per_1000_calls_min = count_close_type/min*1000) %>%
  arrange(desc(per_1000_calls_min))

montgomery_2022_overdoses_zip_with_acs %>%
  mutate(per_1000_calls_max = count_close_type/max*1000) %>%
  arrange(desc(per_1000_calls_max))

```

