---
title: "lab_09.Rmd"
author: "Kiersten Hacker"
date: "2023-04-13"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## You will need

* A Datawrapper account
* Data in CSV files in the `data` folder

## Load libraries and establish settings

**Task** Create a codeblock and load appropriate packages and settings for this lab.

```{r}
# load libraries here
library(tidyverse)
library(janitor)
```

## Questions

**Q1.** You want to explore the differences between the electoral performances of Dan Cox in 2022 and Larry Hogan in 2018. Except a map is probably not the best way to show that. Let's make a chart instead, and the type is up to you, but the goal is to clearly show the differences between the two.

You MUST do the following things when answering this question:
* Load the md_gov_county.csv and md_gov_county_18.csv files from the data folder
* Join those dataframes, specifying the join column. 
* Add columns to generate percentages for Cox and Hogan.
* Export that joined dataframe to a CSV file using `write_csv`.
* Using that exported CSV, make a chart with Datawrapper. What kind of chart is up to you, so experiment with a few kinds of charts to see what answers the question above in the clearest way. Your chart MUST:
  * Have a title, description and source that tells readers about the data. The title and description should highlight the most newsworthy or important aspect.
  * Easily readable labels and values
* Publish your chart
* Describe your thought process below - why you chose to make the chart you did - and add the URL of your chart.

**A1.Link to chart: https://datawrapper.dwcdn.net/GGH7L/1/. I chose to make a stacked bar graph to easily show the comparison of the percentage of votes for Moore in 2022 with the percentage of votes for Hogan in 2018. This way the counties were not doubled for the two values so the graph could be more condensed, even though it is still spread vertically by the 24 jurisdictions. I sorted the bars from the highest to lowest percentage of votes for Moore, since he won the most recent election to keep the chart more current and newsworthy. By changing the stacked bars to show the percentage values in blue for Moore and red for Hogan it is easy to see the distinction for both especially since the colors are associated with their respective political parties. The bar chart is easy to read and follow with labels for the colors, numbers and counties so viewers can take a glance and get an idea without having to do much work to decipher it.  ** 

```{r}
md_gov_county <- read_csv("data/md_gov_county.csv")
md_gov_county_18 <- read_csv("data/md_gov_county_18.csv")

md_gov_county_18_22 <- md_gov_county %>% left_join(md_gov_county_18, by='GEOID') %>%
  mutate(total_votes_22 = Cox + Moore + Lashar + Wallace + Harding + Sefcik + `Write-ins.x`) %>%
  mutate(pct_moore = Moore/total_votes_22 *100) %>%
  mutate(total_votes_18 = Hogan + Jealous + Quinn + Schlakman + `Write-ins.y`) %>%
  mutate(pct_hogan = Hogan/total_votes_18 *100)

write_csv(md_gov_county_18_22, "data/md_gov_county_18_22.csv")
```

**Q2.** You're interested in seeing how the methods of voting have changed in Maryland since 2018. Using the `md_votes_by_type.csv` that has data from the top of the ticket (governor in 2018 and 2022, president in 2020), make a chart showing the percentage of votes cast in each category - early voting, election day and mail - over the three years. You'll need to make those percentages in R and then export the CSV using `write_csv`, then make the chart in Datawrapper. Have a title, description and source that tells readers about the data. The title and description should highlight the most newsworthy or important aspect.

HINT: In Datawrapper's "Check and Describe" tab, you can click on a column to get the option of "hiding" it in any visualization.

What do you think the story here is? Put the answer below along with the URL of your chart.

**A2. Link to the chart: https://datawrapper.dwcdn.net/E7Mg9/1/. I think the story here is that there was a spike in mail-in voting in 2020, which makes sense due to the pandemic. However, in 2018 and in 2022, mail in voting was the least popular method of voting. I think a story looking deeper into why people do not vote by mail as much or which states have passed restrictions on voting by mail or delving deeper into the general narrative around mail-in voting that spread could lead to an interesting piece.** 

```{r}
md_votes_by_type <- read_csv("data/md_votes_by_type.csv") %>%
  rename(Election_day = `Election Day`)

md_votes_by_type <- md_votes_by_type %>%
  mutate(pct_early = Early/Total *100) %>%
  mutate(pct_election_day = Election_day/Total *100) %>%
  mutate(pct_mail = Mail/Total *100)

write_csv(md_votes_by_type, "data/md_votes_by_type.csv")
```


**Q3.** You want to know more about the geography of 911 overdose calls in Montgomery County, Maryland, in 2022. To do that, you'll make a choropleth map in Datawrapper. You will need to:

* Read the file into R and create columns for the number of calls in each zip code and the percentage of all calls that number represents and then export that CSV to use in your map.
* Search for Maryland zip codes (use the 2020 data) as the base map.
* Upload the CSV file.
* Change the tooltip to show both the number and the percentage (you can put them side by side or put one under the other). Make sure to properly format both numbers to make them readable.
* Add labels to your map that locate prominent places in the county.
* Have a title, description and source that tells readers about the data. The title and description should highlight the most newsworthy or important aspect.
* Make any other changes to data that helps you make a more accurate map.

What do you think the story here is, and what would you want to know more about? Put the answer below along with the URL of your map.

**A3. Link to the map: https://datawrapper.dwcdn.net/kLnVd/1/. The story definitely revolves around the Germantown area zip code and the 20852 zip code which covers areas in Rockville and North Bethesda. These two zip codes had the highest percentage of calls, though the zip code covering Germantown was much higher. I think we could look into what towns more specifically had higher numbers of calls since there are other towns aside from Germantown that are included in the zip code. I would want to know what makes the population in the center county areas with higher percentages of calls different from up county or the other portions. Is the policing different? Are the demographics different? Is income different? What initiatives are being taken to combat drug overdoses in these areas.  ** 

```{r}
montgomery_2022_overdoses <- read_csv("data/montgomery_2022_overdoses.csv")
  
montgomery_2022_overdoses_by_zip <- montgomery_2022_overdoses %>%
  group_by (zip) %>%
  summarise (
    count_close_type = n()
  )

montgomery_2022_overdoses_by_zip <- montgomery_2022_overdoses_by_zip %>%
  mutate (pct_calls = count_close_type/659 *100)

write_csv(montgomery_2022_overdoses_by_zip, "data/montgomery_2022_overdoses_by_zip.csv")
```
